---
title: "Seeds"
author: "Emma Ochsner"
output: html_notebook
---

This Notebook is for getting all the seeds of all the data i did.
With this, it is possible to generate the exact same data as I did with my simulations


For the cluster models:

First for the recombination rate 1e-7 (default rate and used in most of the simulations)

Model 1 - Nrnr
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_model1") #set the working directory

options(scipen = 999) #to have regular decimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "sim_log_nonrandom_notrelated") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_model1_nrnr_cluster.csv", row.names = FALSE)

```

Model 1 - Nrr
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_model1")

options(scipen = 999) #to have regular decimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "sim_log_nonrandom_related") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_model1_nrr_cluster.csv", row.names = FALSE)


```

Model 1 - R
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_model1")

options(scipen = 999) #to have regular decimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "sim_log_random_") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_model1_r_cluster.csv", row.names = FALSE)


```



Matings - 16 individuals, Nrnr
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_matings")

options(scipen = 999) #to have regular decimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "mateChoices_nrnr") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.txt","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_matings_16_nrnr_cluster.csv", row.names = FALSE)

```

Matings - 16 individuals, R
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_matings")

options(scipen = 999) #to have regular decimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "mateChoices_r") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.txt","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_matings_16_r_cluster.csv", row.names = FALSE)

```

Matings - 1000 individuals, Nrnr
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_matings_1000ind")

options(scipen = 999) #to have regular decimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "mateChoices_nrnr") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.txt","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_matings_1000_nrnr_cluster.csv", row.names = FALSE)

```


Matings - 1000 individuals, R
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_matings_1000ind")

options(scipen = 999) #to have regular decimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "mateChoices_r") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.txt","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_matings_1000_r_cluster.csv", row.names = FALSE)

```



Rel - het model, 16 individuals - Nrnr
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_rel_het_16")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "relatedness_heterozygosity_nrnr") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.txt","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_rel_het_16_nrnr_cluster.csv", row.names = FALSE)



```

Rel - het model, 16 individuals - Nrr
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_rel_het_16")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "relatedness_heterozygosity_nrr") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.txt","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_rel_het_16_nrr_cluster.csv", row.names = FALSE)



```

Rel - het model, 16 individuals - R
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_rel_het_16")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "relatedness_heterozygosity_r") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.txt","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_rel_het_16_r_cluster.csv", row.names = FALSE)



```

Rel - het model, 1000 individuals - Nrnr
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_rel_het_1000")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "relatedness_heterozygosity_nrnr") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.txt","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_rel_het_1000_nrnr_cluster.csv", row.names = FALSE)



```



Rel - het model, 1000 individuals - Nrr
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_rel_het_1000")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "relatedness_heterozygosity_nrr") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.txt","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_rel_het_1000_nrr_cluster.csv", row.names = FALSE)



```

Rel - het model, 1000 individuals - R
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_rel_het_1000")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "relatedness_heterozygosity_r") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.txt","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_rel_het_1000_r_cluster.csv", row.names = FALSE)


```



Model 2
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_model2_62_100_r")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "model2_") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.txt","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_model2.csv", row.names = FALSE)


```

Model 2.1
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_model2.1")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "model2.1_reps_") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_model2.1_cluster.csv", row.names = FALSE)



```




Recombination rate of 1e-6

Model 1 - Nrnr (1e-6)
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_1e-6_model1")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "sim_log_nonrandom_notrelated") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_1e-6_model1_nrnr.csv", row.names = FALSE)


```

Model 1 - Nrr (1e-6)
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_1e-6_model1")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "sim_log_nonrandom_related") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_1e-6_model1_nrr.csv", row.names = FALSE)


```

Model 1 - R (1e-6)
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_1e-6_model1")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "sim_log_random") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_1e-6_model1_r.csv", row.names = FALSE)


```


Model 2 (1e-6)
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_1e-6_model2")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "model2_") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.txt","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_1e-6_model2.csv", row.names = FALSE)


```

Recombination rate of 1e-8

Model 1 - Nrnr (1e-8)
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_1e-8_model1")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "sim_log_nonrandom_notrelated") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_1e-8_model1_nrnr.csv", row.names = FALSE)


```


Model 1 - Nrr (1e-8)
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_1e-8_model1")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "sim_log_nonrandom_related") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_1e-8_model1_nrr.csv", row.names = FALSE)


```


Model 1 - R (1e-8)
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_1e-8_model1")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "sim_log_random_") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_1e-8_model1_r.csv", row.names = FALSE)


```



Model 2 (1e-8)
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_1e-8_model2")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "model2_") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.txt","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_1e-8_model2.csv", row.names = FALSE)

```


Recombination rate of 1e-9

Model 1 - Nrnr (1e-9)
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_1e-9_model1")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "sim_log_nonrandom_notrelated") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_1e-9_model1_nrnr.csv", row.names = FALSE)


```


Model 1 - Nrr (1e-9)
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_1e-9_model1")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "sim_log_nonrandom_related") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_1e-9_model1_nrr.csv", row.names = FALSE)


```

Model 1 - R (1e-9)
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_1e-9_model1")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "sim_log_random") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_1e-9_model1_r.csv", row.names = FALSE)


```



Model 2 (1e-9)
```{r}
rm(list=ls(all=TRUE)) #clear R environment

setwd("/tmp/server_data_1e-9_model2")

options(scipen = 999) #to have regular ddecimal formatting of the numbers

#list all files with this specific pattern
files <- list.files(pattern = "model2_") 

#add rows for each file, only 1 columns which contains the seed
results_df <- matrix(NA ,nrow = length(files), ncol = (1))
colnames(results_df) <- "seeds"

#then apply these things to all the files stored in the vector
for(i in 1:length(files)){ 
  
  single_file <- files[i] # this is the name of a single SLiM result file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.txt","",single_file)) #use gsub() to get the value of 'seed' that was saved in the name of the results file
  
  results_df[i,"seeds"] <- get_the_seed # record the seed for this particular results file

}

#Now save the values in a table on the desktop
write.csv(results_df, file = "~/Desktop/seeds_1e-9_model2.csv", row.names = FALSE)

```




