---
title: "Inbreeding management model"
author: "Emma Ochsner"
output: html_notebook
---

# Introduction

This is the R code for my Inbreeding management model. I used SLiM to generate data and R to then analyze it. 


## Generating Data in SLiM with Parameters from R - model 1 with non-random mating selecting for not related

```{r eval=FALSE, include=FALSE}

rm(list=ls(all=TRUE)) #clear R environment

options(scipen = 999) #scientific notation penalty, with this R does not use scientific notation for numbers

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory

reps <- 10
repetition <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)


for(r in 1:reps){
  slim_code_block <- paste0("slim -d repetition=", r, " model1_nonrandom_notrelated.slim")
  #slim_code_block 
  system(slim_code_block)#call the code block for slim in the terminal via bash, system invokes the OS command specified by command.
}



#Here i could soft code the variables of the model, maybe the different management strategies (maybe better in SLiM...?)
#parameters will be hard coded in SLiM

#ALso I have to add a codeblock - so it will send all informations needed to SLiM

```



Analyzing the output files from SLiM
model 1 with non-random mating selecting for not related

```{r}
setwd("~/Documents/GitHub/2502_emma_bsc/Outputs/") #setting the working directory
getwd()  # Prints the current working directory
## all SLiM output files which shiw nonrandom mating with mate choice for least relatedness
slim_out_files <- list.files(pattern = "_nonrandom_notrelated_")

length(slim_out_files)
#install.packages("hierfstat")
library("hierfstat")

## make a dataframe to collect results
## each row in the dataframe corresponds to the results from a single SLiM simulation run

row_numbers <- seq(1, 491, by = 10)
#row_names <- paste(row_numbers, "reps", "seed", sep = ";")

results_df <- matrix(NA ,nrow = length(slim_out_files), ncol = (length(row_numbers) + 2))
colnames(results_df) <- c(row_numbers,"reps","seed")
length(c(row_numbers,"reps","seed"))
dim(results_df)


for(i in 1:length(slim_out_files)){ 
  
  single_file <- slim_out_files[i] ## this is the name of a single SLiM result file
  get_the_reps <- as.numeric(gsub(".*_reps_([0-9]+)_.*", "\\1", single_file))  ## use gsub() to get the value of 'reps' that we saved in the name of the results file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) ## and for the seed....
  
  results_df[i,"reps"] <- get_the_reps ## record the value of 'r' for this particular results file
  #.*_reps_ → Matches everything before _reps_
  #([0-9]+) → Captures one or more digits (the reps number)
  #_.* → Matches everything after the number (so it gets removed)
  #"\\1" → Extracts only the number inside ().
  
  results_df[i,"seed"] <- get_the_seed ## record the seed for this particular results file
  
  # Read the CSV file, ensure that it recognizes the header
  df <- read.csv(single_file, header = TRUE) 
  
  # Ensure the column names are correct
  colnames(df) <- c("cycle", "Heterozygosity")  

  # Match cycle numbers to the matrix column names
  matched_indices <- match(df$cycle, colnames(results_df))

  # Assign heterozygosity values to the correct positions in results_df
  results_df[i, matched_indices] <- df$Heterozygosity  

}

# Calculate column means (excluding the last two columns: "reps" and "seed")
column_means <- colMeans(results_df[, 1:(ncol(results_df)-2)], na.rm = TRUE)
# Create a new row with NA for "reps" and "seed" columns
mean_row <- c(column_means, NA, NA)
# Bind the new row to the bottom of results_df
results_df <- rbind(results_df, mean_row)
# Assign row names 
rownames(results_df)[nrow(results_df)] <- "Mean"

# Print the final matrix with the mean row
print(results_df)



print(results_df)

write.csv(results_df, "Results_nonrandom_notrelated.csv", row.names = F) #write a file with my results


```


Now i have my results stored in a matrix. This makes it easier to analyze the results. 
```{r}

# Extract the cycles (generations) from column names - i dont want the last two entries (seed and reps)
#This will be helpful for the plot
cycle_numbers <- as.numeric(colnames(results_df)[1:(ncol(results_df)-2)]) 

# Extract the heterozygosity values (excluding the last row "Mean" and last two columns "reps" & "seed")
heterozygosity_values <- results_df[1:(nrow(results_df)-1), 1:(ncol(results_df)-2)]

# Extract the mean heterozygosity values from the last row --> Last row, so we can use nrow
mean_values <- results_df[nrow(results_df), 1:(ncol(results_df)-2)]

# Create the plot
matplot(
  cycle_numbers,  # X-axis: Generations
  t(heterozygosity_values),  # Transpose to align data correctly
  type = "p",  # "p" for points
  pch = 16,  # Use small dots for points
  col = rgb(0.4, 0.6, 1, 0.2),  # Blue with transparency --> rgb is kind of nice, (red, green, blue, alpha)
  xlab = "Generation (Cycle)",
  ylab = "Heterozygosity",
  main = "Heterozygosity Across Generations - nonrandom-mating selecting for nonrelatedness",
  cex.main = 0.8  # Makes title smaller
)

# Add the mean line
lines(cycle_numbers, mean_values, col = "orange", lwd = 2)  # Red solid line



```



## Generating Data in SLiM with Parameters from R - model 1 with non-random mating selecting for relatedness

```{r eval=FALSE, include=FALSE}

rm(list=ls(all=TRUE)) #clear R environment

options(scipen = 999) #scientific notation penalty, with this R does not use scientific notation for numbers

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory

reps <- 10
repetition <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)


for(r in 1:reps){
  slim_code_block <- paste0("slim -d repetition=", r, " model1_nonrandom_related.slim")
  #slim_code_block 
  system(slim_code_block)#call the code block for slim in the terminal via bash, system invokes the OS command specified by command.
}

```



Analyzing the output files from SLiM
model 1 with non-random mating selecting for relatedness

```{r}
#setwd("~/Documents/GitHub/2502_emma_bsc/Outputs/") #setting the working directory
getwd()  # Prints the current working directory
## all SLiM output files which shiw nonrandom mating with mate choice for least relatedness
slim_out_files <- list.files(pattern = "_nonrandom_related_")

length(slim_out_files)
#install.packages("hierfstat")
library("hierfstat")

## make a dataframe to collect results
## each row in the dataframe corresponds to the results from a single SLiM simulation run

row_numbers <- seq(1, 491, by = 10) #this will help me create the columns i need in my matrix
#row_names <- paste(row_numbers, "reps", "seed", sep = ";")

results_df <- matrix(NA ,nrow = length(slim_out_files), ncol = (length(row_numbers) + 2))
colnames(results_df) <- c(row_numbers,"reps","seed")
length(c(row_numbers,"reps","seed"))
dim(results_df)


for(i in 1:length(slim_out_files)){ 
  
  single_file <- slim_out_files[i] ## this is the name of a single SLiM result file
  get_the_reps <- as.numeric(gsub(".*_reps_([0-9]+)_.*", "\\1", single_file))  ## use gsub() to get the value of 'reps' that we saved in the name of the results file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) ## and for the seed....
  
  results_df[i,"reps"] <- get_the_reps ## record the value of 'r' for this particular results file
  #.*_reps_ → Matches everything before _reps_
  #([0-9]+) → Captures one or more digits (the reps number)
  #_.* → Matches everything after the number (so it gets removed)
  #"\\1" → Extracts only the number inside ().
  
  results_df[i,"seed"] <- get_the_seed ## record the seed for this particular results file
  
  # Read the CSV file, ensure that it recognizes the header
  df <- read.csv(single_file, header = TRUE) 
  
  # Ensure the column names are correct
  colnames(df) <- c("cycle", "Heterozygosity")  

  # Match cycle numbers to the matrix column names
  matched_indices <- match(df$cycle, colnames(results_df))

  # Assign heterozygosity values to the correct positions in results_df
  results_df[i, matched_indices] <- df$Heterozygosity  

}

# Calculate column means (excluding the last two columns: "reps" and "seed")
column_means <- colMeans(results_df[, 1:(ncol(results_df)-2)], na.rm = TRUE)
# Create a new row with NA for "reps" and "seed" columns
mean_row <- c(column_means, NA, NA)
# Bind the new row to the bottom of results_df
results_df <- rbind(results_df, mean_row)
# Assign row names 
rownames(results_df)[nrow(results_df)] <- "Mean"

# Print the final matrix with the mean row
print(results_df)



print(results_df)

write.csv(results_df, "Results_nonrandom_related.csv", row.names = F) #write a file with my results


```


Now i have my results stored in a matrix. This makes it easier to analyze the results. 
```{r}

# Extract the cycles (generations) from column names - i dont want the last two entries (seed and reps)
#This will be helpful for the plot
cycle_numbers <- as.numeric(colnames(results_df)[1:(ncol(results_df)-2)]) 

# Extract the heterozygosity values (excluding the last row "Mean" and last two columns "reps" & "seed")
heterozygosity_values <- results_df[1:(nrow(results_df)-1), 1:(ncol(results_df)-2)]

# Extract the mean heterozygosity values from the last row --> Last row, so we can use nrow
mean_values <- results_df[nrow(results_df), 1:(ncol(results_df)-2)]

# Create the plot
matplot(
  cycle_numbers,  # X-axis: Generations
  t(heterozygosity_values),  # Transpose to align data correctly
  type = "p",  # "p" for points
  pch = 16,  # Use small dots for points
  col = rgb(0.4, 0.6, 1, 0.2),  # Blue with transparency --> rgb is kind of nice, (red, green, blue, alpha)
  xlab = "Generation (Cycle)",
  ylab = "Heterozygosity",
  main = "Heterozygosity Across Generations - nonrandom-mating selecting for relatedness",
  cex.main = 0.8  # Makes title smaller
)

# Add the mean line
lines(cycle_numbers, mean_values, col = "orange", lwd = 2)  # Red solid line



```



Now i do exactly the same, but without any mate choice command in slim. So the mating will be random...


## Generating Data in SLiM with Parameters from R - model 1 with non-random mating selecting for relatedness

```{r eval=FALSE, include=FALSE}

rm(list=ls(all=TRUE)) #clear R environment

options(scipen = 999) #scientific notation penalty, with this R does not use scientific notation for numbers

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory

reps <- 10
repetition <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)


for(r in 1:reps){
  slim_code_block <- paste0("slim -d repetition=", r, " model1_random.slim")
  #slim_code_block 
  system(slim_code_block)#call the code block for slim in the terminal via bash, system invokes the OS command specified by command.
}

```



Analyzing the output files from SLiM
model 1 with non-random mating selecting for relatedness

```{r}
#setwd("~/Documents/GitHub/2502_emma_bsc/Outputs/") #setting the working directory
getwd()  # Prints the current working directory
## all SLiM output files which shiw nonrandom mating with mate choice for least relatedness
slim_out_files <- list.files(pattern = "_random_")

length(slim_out_files)
#install.packages("hierfstat")
library("hierfstat")

## make a dataframe to collect results
## each row in the dataframe corresponds to the results from a single SLiM simulation run

row_numbers <- seq(1, 491, by = 10) #this will help me create the columns i need in my matrix
#row_names <- paste(row_numbers, "reps", "seed", sep = ";")

results_df <- matrix(NA ,nrow = length(slim_out_files), ncol = (length(row_numbers) + 2))
colnames(results_df) <- c(row_numbers,"reps","seed")
length(c(row_numbers,"reps","seed"))
dim(results_df)


for(i in 1:length(slim_out_files)){ 
  
  single_file <- slim_out_files[i] ## this is the name of a single SLiM result file
  get_the_reps <- as.numeric(gsub(".*_reps_([0-9]+)_.*", "\\1", single_file))  ## use gsub() to get the value of 'reps' that we saved in the name of the results file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) ## and for the seed....
  
  results_df[i,"reps"] <- get_the_reps ## record the value of 'r' for this particular results file
  #.*_reps_ → Matches everything before _reps_
  #([0-9]+) → Captures one or more digits (the reps number)
  #_.* → Matches everything after the number (so it gets removed)
  #"\\1" → Extracts only the number inside ().
  
  results_df[i,"seed"] <- get_the_seed ## record the seed for this particular results file
  
  # Read the CSV file, ensure that it recognizes the header
  df <- read.csv(single_file, header = TRUE) 
  
  # Ensure the column names are correct
  colnames(df) <- c("cycle", "Heterozygosity")  

  # Match cycle numbers to the matrix column names
  matched_indices <- match(df$cycle, colnames(results_df))

  # Assign heterozygosity values to the correct positions in results_df
  results_df[i, matched_indices] <- df$Heterozygosity  

}

# Calculate column means (excluding the last two columns: "reps" and "seed")
column_means <- colMeans(results_df[, 1:(ncol(results_df)-2)], na.rm = TRUE)
# Create a new row with NA for "reps" and "seed" columns
mean_row <- c(column_means, NA, NA)
# Bind the new row to the bottom of results_df
results_df <- rbind(results_df, mean_row)
# Assign row names 
rownames(results_df)[nrow(results_df)] <- "Mean"

# Print the final matrix with the mean row
print(results_df)



print(results_df)

write.csv(results_df, "Results_random.csv", row.names = F) #write a file with my results


```




Now i have my results stored in a matrix. This makes it easier to analyze the results. 
```{r}

# Extract the cycles (generations) from column names - i dont want the last two entries (seed and reps)
#This will be helpful for the plot
cycle_numbers <- as.numeric(colnames(results_df)[1:(ncol(results_df)-2)]) 

# Extract the heterozygosity values (excluding the last row "Mean" and last two columns "reps" & "seed")
heterozygosity_values <- results_df[1:(nrow(results_df)-1), 1:(ncol(results_df)-2)]

# Extract the mean heterozygosity values from the last row --> Last row, so we can use nrow
mean_values <- results_df[nrow(results_df), 1:(ncol(results_df)-2)]

# Create the plot
matplot(
  cycle_numbers,  # X-axis: Generations
  t(heterozygosity_values),  # Transpose to align data correctly
  type = "p",  # "p" for points
  pch = 16,  # Use small dots for points
  col = rgb(0.4, 0.6, 1, 0.2),  # Blue with transparency --> rgb is kind of nice, (red, green, blue, alpha)
  xlab = "Generation (Cycle)",
  ylab = "Heterozygosity",
  main = "Heterozygosity Across Generations - random-mating",
  cex.main = 0.8  # Makes title smaller
)

# Add the mean line
lines(cycle_numbers, mean_values, col = "orange", lwd = 2)  # Red solid line



```


now i want to combine all the three graphs i just generated into one. This helps me to better compare the outputs...

```{r}



```


