---
title: "Inbreeding management model"
author: "Emma Ochsner"
output: html_notebook
---

This notebook is building on the other notebook (R_notebook_bsc_model1) - here i want to focus on the firs 100 generations.
To generate the data, i changed some parameters in SLiM: The generations and especially also the amount of outputs generated. Now I want every second cycle to print out the heterozygosity.

```{r eval=FALSE, include=FALSE}

rm(list=ls(all=TRUE)) #clear R environment

options(scipen = 999) #scientific notation penalty, with this R does not use scientific notation for numbers

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory

reps <- 20
repetition <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20)


for(r in 1:reps){
  slim_code_block <- paste0("slim -d repetition=", r, " model1_nonrandom_notrelated.slim")
  #slim_code_block 
  system(slim_code_block)#call the code block for slim in the terminal via bash, system invokes the OS command specified by command.
}

```


```{r eval=FALSE, include=FALSE}

rm(list=ls(all=TRUE)) #clear R environment

options(scipen = 999) #scientific notation penalty, with this R does not use scientific notation for numbers

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory

reps <- 20
repetition <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20)


for(r in 1:reps){
  slim_code_block <- paste0("slim -d repetition=", r, " model1_nonrandom_related.slim")
  #slim_code_block 
  system(slim_code_block)#call the code block for slim in the terminal via bash, system invokes the OS command specified by command.
}

```


```{r eval=FALSE, include=FALSE}

rm(list=ls(all=TRUE)) #clear R environment

options(scipen = 999) #scientific notation penalty, with this R does not use scientific notation for numbers

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory

reps <- 20
repetition <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20)


for(r in 1:reps){
  slim_code_block <- paste0("slim -d repetition=", r, " model1_random.slim")
  #slim_code_block 
  system(slim_code_block)#call the code block for slim in the terminal via bash, system invokes the OS command specified by command.
}

```



now i want to analyze the data and make one single plot.

```{r}
slim_out_files <- list.files(pattern = "_nonrandom_notrelated_")

length(slim_out_files)
#install.packages("hierfstat")
library("hierfstat")

## make a dataframe to collect results
## each row in the dataframe corresponds to the results from a single SLiM simulation run

row_numbers <- seq(100001, 100050, by = 2)

results_df_nrnr <- matrix(NA ,nrow = length(slim_out_files), ncol = (length(row_numbers) + 2))
colnames(results_df_nrnr) <- c(row_numbers,"reps","seed")
length(c(row_numbers,"reps","seed"))
dim(results_df_nrnr)


for(i in 1:length(slim_out_files)){ 
  
  single_file <- slim_out_files[i] ## this is the name of a single SLiM result file
  get_the_reps <- as.numeric(gsub(".*_reps_([0-9]+)_.*", "\\1", single_file))  ## use gsub() to get the value of 'reps' that we saved in the name of the results file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) ## and for the seed....
  
  results_df_nrnr[i,"reps"] <- get_the_reps ## record the value of 'r' for this particular results file
  #.*_reps_ → Matches everything before _reps_
  #([0-9]+) → Captures one or more digits (the reps number)
  #_.* → Matches everything after the number (so it gets removed)
  #"\\1" → Extracts only the number inside ().
  
  results_df_nrnr[i,"seed"] <- get_the_seed ## record the seed for this particular results file
  
  # Read the CSV file, ensure that it recognizes the header
  df <- read.csv(single_file, header = TRUE) 
  
  # Ensure the column names are correct
  colnames(df) <- c("cycle", "Heterozygosity")  

  # Match cycle numbers to the matrix column names
  matched_indices <- match(as.character(df$cycle), colnames(results_df_nrnr))

  # Assign heterozygosity values to the correct positions in results_df_nrnr
  results_df_nrnr[i, matched_indices] <- df$Heterozygosity  

}

# Calculate column means (excluding the last two columns: "reps" and "seed")
column_means <- colMeans(results_df_nrnr[, 1:(ncol(results_df_nrnr)-2)], na.rm = TRUE)
# Create a new row with NA for "reps" and "seed" columns
mean_row <- c(column_means, NA, NA)
# Bind the new row to the bottom of results_df_nrnr
results_df_nrnr <- rbind(results_df_nrnr, mean_row)
# Assign row names 
rownames(results_df_nrnr)[nrow(results_df_nrnr)] <- "Mean"

# Print the final matrix with the mean row
print(results_df_nrnr)

write.csv(results_df_nrnr, "Results_nonrandom_notrelated.csv", row.names = F) #write a file with my results


```


Analyzing the output files from SLiM
model 1 with non-random mating selecting for relatedness

```{r}
#setwd("~/Documents/GitHub/2502_emma_bsc/Outputs/") #setting the working directory
getwd()  # Prints the current working directory
## all SLiM output files which shiw nonrandom mating with mate choice for least relatedness
slim_out_files <- list.files(pattern = "_nonrandom_related_")

length(slim_out_files)
#install.packages("hierfstat")
library("hierfstat")

## make a dataframe to collect results
## each row in the dataframe corresponds to the results from a single SLiM simulation run

row_numbers <- seq(100001, 100050, by = 2)
#row_names <- paste(row_numbers, "reps", "seed", sep = ";")

results_df_nrr <- matrix(NA ,nrow = length(slim_out_files), ncol = (length(row_numbers) + 2))
colnames(results_df_nrr) <- c(row_numbers,"reps","seed")
length(c(row_numbers,"reps","seed"))
dim(results_df_nrr)


for(i in 1:length(slim_out_files)){ 
  
  single_file <- slim_out_files[i] ## this is the name of a single SLiM result file
  get_the_reps <- as.numeric(gsub(".*_reps_([0-9]+)_.*", "\\1", single_file))  ## use gsub() to get the value of 'reps' that we saved in the name of the results file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) ## and for the seed....
  
  results_df_nrr[i,"reps"] <- get_the_reps ## record the value of 'r' for this particular results file
  #.*_reps_ → Matches everything before _reps_
  #([0-9]+) → Captures one or more digits (the reps number)
  #_.* → Matches everything after the number (so it gets removed)
  #"\\1" → Extracts only the number inside ().
  
  results_df_nrr[i,"seed"] <- get_the_seed ## record the seed for this particular results file
  
  # Read the CSV file, ensure that it recognizes the header
  df <- read.csv(single_file, header = TRUE) 
  
  # Ensure the column names are correct
  colnames(df) <- c("cycle", "Heterozygosity")  

  # Match cycle numbers to the matrix column names
  matched_indices <- match(df$cycle, colnames(results_df_nrr))

  # Assign heterozygosity values to the correct positions in results_df_nrr
  results_df_nrr[i, matched_indices] <- df$Heterozygosity  

}

# Calculate column means (excluding the last two columns: "reps" and "seed")
column_means <- colMeans(results_df_nrr[, 1:(ncol(results_df_nrr)-2)], na.rm = TRUE)
# Create a new row with NA for "reps" and "seed" columns
mean_row <- c(column_means, NA, NA)
# Bind the new row to the bottom of results_df_nrr
results_df_nrr <- rbind(results_df_nrr, mean_row)
# Assign row names 
rownames(results_df_nrr)[nrow(results_df_nrr)] <- "Mean"

# Print the final matrix with the mean row
print(results_df_nrr)

write.csv(results_df_nrr, "Results_nonrandom_related.csv", row.names = F) #write a file with my results


```


Analyzing the output files from SLiM
model 1 with random mating 

```{r}
#setwd("~/Documents/GitHub/2502_emma_bsc/Outputs/") #setting the working directory
getwd()  # Prints the current working directory
## all SLiM output files which shiw nonrandom mating with mate choice for least relatedness
slim_out_files <- list.files(pattern = "_random_")

length(slim_out_files)
#install.packages("hierfstat")
library("hierfstat")

## make a dataframe to collect results
## each row in the dataframe corresponds to the results from a single SLiM simulation run

row_numbers <- seq(100001, 100050, by = 2)
#row_names <- paste(row_numbers, "reps", "seed", sep = ";")

results_df_r <- matrix(NA ,nrow = length(slim_out_files), ncol = (length(row_numbers) + 2))
colnames(results_df_r) <- c(row_numbers,"reps","seed")
length(c(row_numbers,"reps","seed"))
dim(results_df_r)


for(i in 1:length(slim_out_files)){ 
  
  single_file <- slim_out_files[i] ## this is the name of a single SLiM result file
  get_the_reps <- as.numeric(gsub(".*_reps_([0-9]+)_.*", "\\1", single_file))  ## use gsub() to get the value of 'reps' that we saved in the name of the results file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) ## and for the seed....
  
  results_df_r[i,"reps"] <- get_the_reps ## record the value of 'r' for this particular results file
  #.*_reps_ → Matches everything before _reps_
  #([0-9]+) → Captures one or more digits (the reps number)
  #_.* → Matches everything after the number (so it gets removed)
  #"\\1" → Extracts only the number inside ().
  
  results_df_r[i,"seed"] <- get_the_seed ## record the seed for this particular results file
  
  # Read the CSV file, ensure that it recognizes the header
  df <- read.csv(single_file, header = TRUE) 
  
  # Ensure the column names are correct
  colnames(df) <- c("cycle", "Heterozygosity")  

  # Match cycle numbers to the matrix column names
  matched_indices <- match(df$cycle, colnames(results_df_r))

  # Assign heterozygosity values to the correct positions in results_df_r
  results_df_r[i, matched_indices] <- df$Heterozygosity  

}

# Calculate column means (excluding the last two columns: "reps" and "seed")
column_means <- colMeans(results_df_r[, 1:(ncol(results_df_r)-2)], na.rm = TRUE)
# Create a new row with NA for "reps" and "seed" columns
mean_row <- c(column_means, NA, NA)
# Bind the new row to the bottom of results_df_r
results_df_r <- rbind(results_df_r, mean_row)
# Assign row names 
rownames(results_df_r)[nrow(results_df_r)] <- "Mean"

# Print the final matrix with the mean row
print(results_df_r)

write.csv(results_df_r, "Results_random.csv", row.names = F) #write a file with my results


```


Now do the plot:
```{r}

# Extract the cycles (generations) from column names - i dont want the last two entries (seed and reps)
#This will be helpful for the plot
cycle_numbers <- as.numeric(colnames(results_df_nrnr)[1:(ncol(results_df_nrnr)-2)]) 
# Extract the heterozygosity values (excluding the last row "Mean" and last two columns "reps" & "seed")
heterozygosity_values_nrnr <- results_df_nrnr[1:(nrow(results_df_nrnr)-1), 1:(ncol(results_df_nrnr)-2)]
# Extract the mean heterozygosity values from the last row --> Last row, so we can use nrow
mean_values_nrnr <- results_df_nrnr[nrow(results_df_nrnr), 1:(ncol(results_df_nrnr)-2)]


# Extract the heterozygosity values (excluding the last row "Mean" and last two columns "reps" & "seed")
heterozygosity_values_nrr <- results_df_nrr[1:(nrow(results_df_nrr)-1), 1:(ncol(results_df_nrr)-2)]
# Extract the mean heterozygosity values from the last row --> Last row, so we can use nrow
mean_values_nrr <- results_df_nrr[nrow(results_df_nrr), 1:(ncol(results_df_nrr)-2)]

# Extract the heterozygosity values (excluding the last row "Mean" and last two columns "reps" & "seed")
heterozygosity_values_r <- results_df_r[1:(nrow(results_df_r)-1), 1:(ncol(results_df_r)-2)]
# Extract the mean heterozygosity values from the last row --> Last row, so we can use nrow
mean_values_r <- results_df_r[nrow(results_df_r), 1:(ncol(results_df_r)-2)]





heterozygosity_only_nrnr <- heterozygosity_values_nrnr[, 1:length(cycle_numbers)]
heterozygosity_only_nrr <- heterozygosity_values_nrr[, 1:length(cycle_numbers)]
heterozygosity_only_r <- heterozygosity_values_r[, 1:length(cycle_numbers)]
#I did this to cut off the last two columns (seed and rep) --> Due to this, i can then apply the filter (next lines) to the matrix. Otherwise it would not match the dimension of cycle_numbers...


# Create the plot
matplot(
  cycle_numbers,  # X-axis: Generations, here the filter is active 
  t(heterozygosity_only_r),  # Transpose to align data correctly --> I want to cut the columns not the rows, thats why there is this notation with the ,
  type = "p",  # "p" for points
  pch = 16,  # Use small dots for points
  col = rgb(0.4, 0.6, 1, 0.2),  # light Blue with transparency --> rgb is kind of nice, (red, green, blue, alpha)
  xlab = "Generation (Cycle)",
  ylab = "Heterozygosity (log)",
  main = "Heterozygosity Across Generations",
  cex.main = 0.8,  # Makes title smaller
  log = "y"  # This makes the y-axis logarithmic
)

# Add the mean line
lines(cycle_numbers, mean_values_r, col = "blue", lwd = 2)  # Red solid line

# Now add the other points and lines
# Add non-random selecting for relatedness
matpoints(cycle_numbers, t(heterozygosity_only_nrr), pch = 16, col = rgb(1, 0.5, 0.5, 0.2))
lines(cycle_numbers, mean_values_nrr, col = "red", lwd = 2)

# add non-rendom selecting for not related
matpoints(cycle_numbers, t(heterozygosity_only_nrnr), pch = 16, col = rgb(0.4, 1, 0.4, 0.2))
lines(cycle_numbers, mean_values_nrnr, col = "forestgreen", lwd = 2)

legend("topright", legend = c("Random mating", "Non-random mating selecting for relatedness", "Non-random mating selecting for non-relatedness"), col = c("blue", "red", "forestgreen"), lwd = 2, bty = "n", cex = 0.6)


```





