---
title: "R Notebook"
output: html_notebook
---

This notebook helps looking at the rates of decline in heterozygosity and increase in relatedness for the different mateCHoice models.

I have simulated the data on the server and want to analyze it now. For this i've created a folder, which is connected to the output server of this model. In this way, i dont have to load all the data on my computer.

The size of the starting population is 1000 individuals or 16 individuals, which have different genomes and they mate for 100 generations.  

```{r}

#This is for the output on the server for the model with 16 incividuals
setwd("/tmp/server_data_rel_het_16")
#This is for the output on the server for the model with 1000 incividuals
#setwd("/tmp/server_data_rel_het_1000")

rm(list=ls(all=TRUE)) #clear R environment

files_nrnr <- list.files(pattern = "_heterozygosity_nrnr_") #list all files with this pattern in the name
length(files_nrnr) #check if it got all the files

#This will create a list of data frames, where every element is a replicate. Each replicate will have the column: Generation, Relatedness and heterozygosity. 
rep_data_nrnr <- lapply(files_nrnr, function(f) {
  read.table(f, col.names = c("generation", "relatedness", "heterozygosity"))
})
#lapply is a command which applies a function to all the elements in a vector - So here it names the columns of every file, stored in files_nrnr

# Assume all replicates have the same generations - define the generations for later plot them
generations <- rep_data_nrnr[[1]]$generation

# Extract relatedness into a new matrix: each column = one replicate 
rel_matrix_nrnr <- sapply(rep_data_nrnr, function(df) df$relatedness)
#Specifically this tells R: Go through every element "df" in the list rep_data_nrnr, and extract the relatedness column. 

# Compute mean relatedness per generation
rel_mean_nrnr <- rowMeans(rel_matrix_nrnr)

#Now do the same for heterozygosity
het_matrix_nrnr <- sapply(rep_data_nrnr, function(df) df$heterozygosity) #create a new matrix for all heterozygosity values
het_mean_nrnr <- rowMeans(het_matrix_nrnr)

#And now again all this for the data of the model which has non-random mating selecting for relatedness (nrr)

files_nrr <- list.files(pattern = "_heterozygosity_nrr_")
length(files_nrr)

rep_data_nrr <- lapply(files_nrr, function(f) {
  read.table(f, col.names = c("generation", "relatedness", "heterozygosity"))
})

rel_matrix_nrr <- sapply(rep_data_nrr, function(df) df$relatedness)
rel_mean_nrr <- rowMeans(rel_matrix_nrr)

het_matrix_nrr <- sapply(rep_data_nrr, function(df) df$heterozygosity)
het_mean_nrr <- rowMeans(het_matrix_nrr)


#And now again all this for the data of the model which has random mating (r)

files_r <- list.files(pattern = "_heterozygosity_r_")
length(files_r)

rep_data_r <- lapply(files_r, function(f) {
  read.table(f, col.names = c("generation", "relatedness", "heterozygosity"))
}) #lapply: apply a function over a list 

rel_matrix_r <- sapply(rep_data_r, function(df) df$relatedness)
rel_mean_r <- rowMeans(rel_matrix_r)

het_matrix_r <- sapply(rep_data_r, function(df) df$heterozygosity)
het_mean_r <- rowMeans(het_matrix_r)

```


Now i have sorted my data and can combine them into 2 graps. I do one graph which shows the relatedness over the generations in the 3 different models and one for the heterozygosity values. 
```{r}


par(mfrow = c(1, 2)) #To have 2 plots shown at once
par(mfrow = c(1, 2), oma = c(2, 0, 1, 0))  # oma = outer margin area (bottom, left, top, right), leaves space for a title and some space at the bottom to add a legend outside of the plots. 

par(mgp = c(1.8, 0.7, 0)) #This is to decrease the distance of the axis labels to the axis. 

par(cex.lab = 0.8, cex.axis = 0.7, cex.main = 1) #this helps me rescaling the size of the labels

# Combine both matrices into one big matrix to find global max and min
max_rel <- max(rel_matrix_nrnr, rel_matrix_nrr, rel_matrix_r)
min_rel <- min(rel_matrix_nrnr, rel_matrix_nrr, rel_matrix_r)

x_vals <- 1:100

# Model nrnr: light blue points
#Here i dont have to transpose the data, since each column is a separate data series
matplot(x_vals, rel_matrix_nrnr, type = "p", pch = 16, cex = 0.3,
        col = adjustcolor("#54BFB7FF", alpha.f = 0.01),
        xlab = "Generation", ylab = "Relatedness",
        ylim = c(min_rel, max_rel))  


# Model nrr: light green points
matpoints(x_vals, rel_matrix_nrr, pch = 16, cex = 0.3, col = adjustcolor("#894B33FF", alpha.f = 0.01)) #adding points to the existing plot from above

# Model r: light purple points
matpoints(x_vals, rel_matrix_r, pch = 16, cex = 0.2, col = adjustcolor("tan1", alpha.f = 0.01)) #again

# === adding Mean relatedness lines to the plot ===
lines(x_vals, rel_mean_nrnr, col = "#54BFB7FF", lwd = 2.5)
lines(x_vals, rel_mean_nrr, col = "#894B33FF", lwd = 2)
lines(x_vals, rel_mean_r, col = "tan1", lwd = 2)


#And now the exactly same but for the heterozygosity values

# Combine both matrices into one big matrix to find global max and min
max_het <- max(het_matrix_nrnr, het_matrix_nrr, het_matrix_r)
min_het <- min(het_matrix_nrnr, het_matrix_nrr, het_matrix_r)

# Model nrnr: light blue points
matplot(x_vals, het_matrix_nrnr, type = "p", pch = 16, cex = 0.3,
        col = adjustcolor("#54BFB7FF", alpha.f = 0.01),
        xlab = "Generation", ylab = "Heterozygosity", 
        ylim = c(min_het, max_het))  


# Model nrr: light green points
matpoints(x_vals, het_matrix_nrr, pch = 16, cex = 0.3, col = adjustcolor("#894B33FF", alpha.f = 0.01)) #add points

# Model r: light purple points
matpoints(x_vals, het_matrix_r, pch = 16, cex = 0.2, col = adjustcolor("tan1", alpha.f = 0.01)) #add points

# === adding Mean relatedness lines to the plot ===
lines(x_vals, het_mean_nrnr, col = "#54BFB7FF", lwd = 2)
lines(x_vals, het_mean_nrr, col = "#894B33FF", lwd = 2)
lines(x_vals, het_mean_r, col = "tan1", lwd = 2)

mtext("Relatedness and Heterozygosity Over Time", side = 3, line = -2, outer = TRUE, cex = 1.2) #creates a shared title

par(xpd = NA) #Allows drawing outside of the plot - this is for the legend
legend("bottom", inset = c(0, -0.6), horiz = FALSE,
       legend = c("not-random not-related", "not-random related", "random"),
       col = c("#54BFB7FF", "#894B33FF", "tan1"),
       lwd = 2, bty = "n", cex = 0.6)
#inset: Pushes it down below the plot

```

And again the same graphs but without the points for better visualization...

```{r}


par(mfrow = c(1, 2)) #To have 2 plots shown at once
par(mfrow = c(1, 2), oma = c(2, 0, 1, 0))  # oma = outer margin area (bottom, left, top, right), leaves space for a title and some space at the bottom to add a legend outside of the plots. 

par(mgp = c(1.8, 0.7, 0)) #This is to decrease the distance of the axis labels to the axis. 

par(cex.lab = 0.8, cex.axis = 0.7, cex.main = 1) #this helps me rescaling the size of the labels

# Combine both matrices into one big matrix to find global max and min
max_rel <- max(rel_matrix_nrnr, rel_matrix_nrr, rel_matrix_r)
min_rel <- min(rel_matrix_nrnr, rel_matrix_nrr, rel_matrix_r)

x_vals <- 1:100

# Model nrnr: light blue points
#Here i dont have to transpose the data, since each column is a separate data series
matplot(x_vals, rel_mean_nrnr, col = "#54BFB7FF", lwd = 2.5, lty = 1, type = "n",
        xlab = "Generation", ylab = "Relatedness",
        ylim = c(min_rel, max_rel))  



# === adding Mean relatedness lines to the plot ===
lines(x_vals, rel_mean_nrnr, col = "#54BFB7FF", lwd = 2.5)
lines(x_vals, rel_mean_nrr, col = "#894B33FF", lwd = 2)
lines(x_vals, rel_mean_r, col = "tan1", lwd = 2, lty = 4)


#And now the exactly same but for the heterozygosity values

# Combine both matrices into one big matrix to find global max and min
max_het <- max(het_matrix_nrnr, het_matrix_nrr, het_matrix_r)
min_het <- min(het_matrix_nrnr, het_matrix_nrr, het_matrix_r)

# Model nrnr: light blue points
matplot(x_vals, het_mean_nrnr, col = "#54BFB7FF", lwd = 2, lty = 1, type = "n",
        xlab = "Generation", ylab = "Heterozygosity", 
        ylim = c(min_het, max_het))  


# === adding Mean relatedness lines to the plot ===
lines(x_vals, het_mean_nrnr, col = "#54BFB7FF", lwd = 2)
lines(x_vals, het_mean_nrr, col = "#894B33FF", lwd = 2)
lines(x_vals, het_mean_r, col = "tan1", lwd = 2, lty = 4)

mtext("Relatedness and Heterozygosity Over Time", side = 3, line = -2, outer = TRUE, cex = 1.2) #creates a shared title

par(xpd = NA) #Allows drawing outside of the plot - this is for the legend
legend("bottom", inset = c(0, -0.6), horiz = FALSE,
       legend = c("not-random not-related", "not-random related", "random"),
       col = c("#54BFB7FF", "#894B33FF", "tan1"),
       lwd = 2, bty = "n", cex = 0.6)
#inset: Pushes it down below the plot

```

Maybe also one where one can see median and 25/75%

```{r}

par(mfrow = c(1, 2)) #To have 2 plots shown at once
par(mfrow = c(1, 2), oma = c(2, 0, 1, 0))  # oma = outer margin area (bottom, left, top, right), leaves space for a title and some space at the bottom to add a legend outside of the plots. 

par(mgp = c(1.8, 0.7, 0)) #This is to decrease the distance of the axis labels to the axis. 

par(cex.lab = 0.8, cex.axis = 0.7, cex.main = 1) #this helps me rescaling the size of the labels

#Compute median & quartiles (25 and 75 %)
#the 1 means i want to apply the functions over the rows of my table and ignoring NA values (na.rm = TRUE)
#first for the heterozygosity values
het_med_nrnr <- apply(het_matrix_nrnr, 1, median, na.rm = TRUE)
het_p25_nrnr <- apply(het_matrix_nrnr, 1, function(x) quantile(x, 0.25, na.rm = TRUE))
het_p75_nrnr <- apply(het_matrix_nrnr, 1, function(x) quantile(x, 0.75, na.rm = TRUE))

het_med_nrr  <- apply(het_matrix_nrr,  1, median, na.rm = TRUE)
het_p25_nrr  <- apply(het_matrix_nrr,  1, function(x) quantile(x, 0.25, na.rm = TRUE))
het_p75_nrr  <- apply(het_matrix_nrr,  1, function(x) quantile(x, 0.75, na.rm = TRUE))

het_med_r    <- apply(het_matrix_r,    1, median, na.rm = TRUE)
het_p25_r    <- apply(het_matrix_r,    1, function(x) quantile(x, 0.25, na.rm = TRUE))
het_p75_r    <- apply(het_matrix_r,    1, function(x) quantile(x, 0.75, na.rm = TRUE))

#And now for relatedness values
rel_med_nrnr <- apply(rel_matrix_nrnr, 1, median, na.rm = TRUE)
rel_p25_nrnr <- apply(rel_matrix_nrnr, 1, function(x) quantile(x, 0.25, na.rm = TRUE))
rel_p75_nrnr <- apply(rel_matrix_nrnr, 1, function(x) quantile(x, 0.75, na.rm = TRUE))

rel_med_nrr  <- apply(rel_matrix_nrr,  1, median, na.rm = TRUE)
rel_p25_nrr  <- apply(rel_matrix_nrr,  1, function(x) quantile(x, 0.25, na.rm = TRUE))
rel_p75_nrr  <- apply(rel_matrix_nrr,  1, function(x) quantile(x, 0.75, na.rm = TRUE))

rel_med_r    <- apply(rel_matrix_r,    1, median, na.rm = TRUE)
rel_p25_r    <- apply(rel_matrix_r,    1, function(x) quantile(x, 0.25, na.rm = TRUE))
rel_p75_r    <- apply(rel_matrix_r,    1, function(x) quantile(x, 0.75, na.rm = TRUE))

# Create a vector for the x-values (generations) - this will be used for the x axis of the plot. 
x_vals <- 1:100


#For rel values
#define max and min values
max_rel <- max(rel_matrix_nrnr, rel_matrix_nrr, rel_matrix_r)
min_rel <- min(rel_matrix_nrnr, rel_matrix_nrr, rel_matrix_r)

plot(x_vals, rel_med_nrnr, type = "n",
     xlab = "Generation", ylab = "Relatedness",
     ylim = c(min_rel, max_rel),
     main = "Heterozygosity over Generations \n(median & 25/75%)")

#Draw median (solid line) and quartiles (dashed lines)
# Model: non-random non-related
lines(x_vals, rel_med_nrnr, col = "#54BFB7FF", lwd = 2.5, lty = 1) #lwd is line width, lty is line type
lines(x_vals, rel_p25_nrnr, col = "#54BFB7FF", lwd = 1.5, lty = 3)
lines(x_vals, rel_p75_nrnr, col = "#54BFB7FF", lwd = 1.5, lty = 3)

# Model: non-random related
lines(x_vals, rel_med_nrr, col = "#894B33FF", lwd = 2.5, lty = 1)
lines(x_vals, rel_p25_nrr, col = "#894B33FF", lwd = 1.5, lty = 3)
lines(x_vals, rel_p75_nrr, col = "#894B33FF", lwd = 1.5, lty = 3)

# Model: random
lines(x_vals, rel_med_r, col = "tan1", lwd = 2.5, lty = 1)
lines(x_vals, rel_p25_r, col = "tan1", lwd = 1.5, lty = 3)
lines(x_vals, rel_p75_r, col = "tan1", lwd = 1.5, lty = 3)

#define max and min values
max_het <- max(het_matrix_nrnr, het_matrix_nrr, het_matrix_r)
min_het <- min(het_matrix_nrnr, het_matrix_nrr, het_matrix_r)

#Make a plot with the median and the 25/75% as dashed lines

plot(x_vals, het_med_nrnr, type = "n",
     xlab = "Generation", ylab = "Heterozygosity",
     ylim = c(min_het, max_het),
     main = "Heterozygosity over Generations \n(median & 25/75%)")


#For het values
# Model: non-random non-related
lines(x_vals, het_med_nrnr, col = "#54BFB7FF", lwd = 2.5, lty = 1) #lwd is line width, lty is line type
lines(x_vals, het_p25_nrnr, col = "#54BFB7FF", lwd = 1.5, lty = 3)
lines(x_vals, het_p75_nrnr, col = "#54BFB7FF", lwd = 1.5, lty = 3)

# Model: non-random related
lines(x_vals, het_med_nrr, col = "#894B33FF", lwd = 2.5, lty = 1)
lines(x_vals, het_p25_nrr, col = "#894B33FF", lwd = 1.5, lty = 3)
lines(x_vals, het_p75_nrr, col = "#894B33FF", lwd = 1.5, lty = 3)

# Model: random
lines(x_vals, het_med_r, col = "tan1", lwd = 2.5, lty = 1)
lines(x_vals, het_p25_r, col = "tan1", lwd = 1.5, lty = 3)
lines(x_vals, het_p75_r, col = "tan1", lwd = 1.5, lty = 3)



#Add the legend
par(xpd = NA) #Allows drawing outside of the plot - this is for the legend
legend("bottom", inset = c(0, -0.6), horiz = FALSE, ncol     = 3, 
  legend = c("nrnr median", "nrnr 25/75%", "nrr median", "nrr 25/75%", "r median","r 25/75%"),
       col    = c("#54BFB7FF", "#54BFB7FF",
                  "#894B33FF", "#894B33FF",
                  "tan1",       "tan1"),
       lty    = c(1, 2, 1, 2, 1, 2),
       lwd    = c(2.5, 1.5, 2.5, 1.5, 2.5, 1.5),
      bty = "n", cex = 0.6)
#inset: Pushes it down below the plot


```

