---
title: "Matings - Model 1"
author: "Emma Ochsner"
output: html_notebook
---

First I have 2 chunks, which will send information to my slim model - With this callback I can make 20 repetitions of my simulation in SLiM and save the output files.
I do this for the random mating and for the non-random mating selecting for non-relatedness. 

First for non-random not-related

```{r eval=FALSE, include=FALSE}

rm(list=ls(all=TRUE)) #clear R environment

options(scipen = 999) #scientific notation penalty, with this R does not use scientific notation for numbers

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory

reps <- 20
repetition <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20)


for(r in 1:reps){
  slim_code_block <- paste0("slim -d repetition=", r, " matings_nrnr.slim")
  #slim_code_block 
  system(slim_code_block)#call the code block for slim in the terminal via bash, system invokes the OS command specified by command.
}

```


And now for the random mating

```{r eval=FALSE, include=FALSE}

rm(list=ls(all=TRUE)) #clear R environment

options(scipen = 999) #scientific notation penalty, with this R does not use scientific notation for numbers

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory

reps <- 20
repetition <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20)


for(r in 1:reps){
  slim_code_block <- paste0("slim -d repetition=", r, " matings_r.slim")
  #slim_code_block 
  system(slim_code_block)#call the code block for slim in the terminal via bash, system invokes the OS command specified by command.
}

```

-------

Sort the 20 files (nrnr) to analyze the data

```{r}

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory
#setwd("~/Documents/GitHub/2502_emma_bsc/MATE CHOICES/mateChoices_16-16ind_100gen_20reps")
#setwd("~/Documents/GitHub/2502_emma_bsc/MATE_CHOICES/mateChoices_1000-1000ind_100gen_20reps")
#Here the working directory has to be pointed to where the output files are saved. 

library(dplyr)

# List all files with this specific pattern in a vector
files_nrnr <- list.files(pattern = "mateChoices_nrnr_")

# For each file in the vector, read it in, compute df_counts, and return it
#lappply is a command which applies the following function to all files in the vector
all_reps <- lapply(files_nrnr, function(f) {
  
  # 1. Read the data - create a table to store all the data
  df <- read.table(f, header = FALSE)
  colnames(df) <- c("Generation", "FemaleIndex", "MaleIndex") #name the columns in the table
  
  # 2. Summarize by generation
  #take care - either make 8 or 500 here, depending if we have 16 or 1000 individuals   !!!!!!
  df_counts <- df %>%
    group_by(Generation) %>%
    summarise(
      distinctFemales = n_distinct(FemaleIndex),
      distinctMales   = n_distinct(MaleIndex),
      missingFemales  = 500 - n_distinct(FemaleIndex),
      missingMales    = 500 - n_distinct(MaleIndex),
      # calculate the number of individuals that did not mate in each generation --> compute number of distinct ids and then substract it from the total amount of females / males. 
    )
  
  # Return the summarized table
  df_counts
})

# Bind all replicate summaries into one data frame, with a new column "rep" indicating which replicate
df_all <- bind_rows(all_reps, .id = "rep") 
# Now df_all has columns: rep, Generation, distinctFemales, distinctMales, missingFemales, missingMales

#I now want to group the data by generations, to average the data across all replicates for each generation. 
df_means <- df_all %>%
  group_by(Generation) %>%
  summarise(
    meanMissingFemales = mean(missingFemales),
    meanMissingMales   = mean(missingMales),
    .groups = "drop" #remove the grouping again
  )


```


And make the plot
```{r}

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory
#setwd("~/Documents/GitHub/2502_emma_bsc/MATE CHOICES/mateChoices_16-16ind_100gen_20reps")
#setwd("~/Documents/GitHub/2502_emma_bsc/MATE_CHOICES/mateChoices_1000-1000ind_100gen_20reps")

#compute max and min values over the whole data
max_nrnr <- max(df_means$meanMissingFemales, df_means$meanMissingMales)
min_nrnr <- min(df_means$meanMissingFemales, df_means$meanMissingMales)
  
#define x and y values - this will help me to make the plot
x_vals <- df_means$Generation
y_vals <- df_means$meanMissingFemales

#make the plot
matplot(
  x = x_vals,
  y = y_vals,
  type = "l",           # "b" means points connected by lines - "l" is just a line
  xlab = "Generation",
  ylab = "Number Missing", 
  ylim = c(min_nrnr, max_nrnr), #define range of the y axis
  col = "#54BFB7FF", 
  main = "number of females and males which don't get to mate out of 500") #title of the plot

#now add a line for the males
lines(x_vals, df_means$meanMissingMales, type = "l", col = "tan1")

#add the legend
legend("topright",legend = c("Missing Females", "Missing Males"),
  lty = 1, col = c("#54BFB7FF", "tan1"), cex = 0.6)

```



Sort the 20 files (r) to analyze the data - exactly the same procedure as before

```{r}

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory
#setwd("~/Documents/GitHub/2502_emma_bsc/MATE CHOICES/mateChoices_16-16ind_100gen_20reps")
#setwd("~/Documents/GitHub/2502_emma_bsc/MATE_CHOICES/mateChoices_1000-1000ind_100gen_20reps")

library(dplyr)

# Vector of file names (adjust as needed for your naming/paths)
files_r <- list.files(pattern = "mateChoices_r_")

# For each file, read it in, compute df_counts, and return it
all_reps <- lapply(files_r, function(f) {
  
  # 1. Read the data
  df <- read.table(f, header = FALSE)
  colnames(df) <- c("Generation", "FemaleIndex", "MaleIndex")
  
  # 2. Summarize by generation
  #take care - either make 8 or 500 here, depending if we have 16 or 1000 individuals!!

  df_counts <- df %>%
    group_by(Generation) %>%
    summarise(
      distinctFemales = n_distinct(FemaleIndex),
      distinctMales   = n_distinct(MaleIndex),
      missingFemales  = 500 - n_distinct(FemaleIndex),
      missingMales    = 500 - n_distinct(MaleIndex),
      #.groups = "drop"
    )
  
  # Return the summarized table
  df_counts
})

# Bind all replicate summaries into one data frame, with a new column "rep" indicating which replicate
df_all <- bind_rows(all_reps, .id = "rep") 
# Now df_all has columns: rep, Generation, distinctFemales, distinctMales, missingFemales, missingMales

#I now want to group the data by generations, to average the data across all replicates for each generation. 
df_means <- df_all %>%
  group_by(Generation) %>%
  summarise(
    meanMissingFemales = mean(missingFemales),
    meanMissingMales   = mean(missingMales),
    .groups = "drop" #remove the grouping again
  )


```


And make the plot
```{r}

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory
#setwd("~/Documents/GitHub/2502_emma_bsc/MATE CHOICES/mateChoices_16-16ind_100gen_20reps")
#setwd("~/Documents/GitHub/2502_emma_bsc/MATE_CHOICES/mateChoices_1000-1000ind_100gen_20reps")

max_r <- max(df_means$meanMissingFemales, df_means$meanMissingMales)
min_r <- min(df_means$meanMissingFemales, df_means$meanMissingMales)

  
x_vals <- df_means$Generation
y_vals <- df_means$meanMissingFemales

matplot(
  x = x_vals,
  y = y_vals,
  type = "l",           # "b" means points connected by lines
  xlab = "Generation",
  ylab = "Number Missing", 
  ylim = c(min_r, max_r),
  col = "#54BFB7FF",
    main = "number of females and males which don't get to mate out of 500")

lines(x_vals, df_means$meanMissingMales, type = "l", col = "tan1")

legend("bottomleft",legend = c("Missing Females", "Missing Males"),
  lty = 1, col = c("#54BFB7FF", "tan1"), cex = 0.6)

```



