---
title: "Inbreeding management model"
author: "Emma Ochsner"
output: html_notebook
---

This notebook is building on the other notebook (R_notebook_bsc_model1) - here i want to focus on the firs 100 generations.
To generate the data, i changed some parameters in SLiM: The generations and especially also the amount of outputs generated. Now I want every second cycle to print out the heterozygosity.

If the model was run on the server - do not use this command
```{r eval=FALSE, include=FALSE}

rm(list=ls(all=TRUE)) #clear R environment

options(scipen = 999) #scientific notation penalty, with this R does not use scientific notation for numbers

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory

reps <- 50
repetition <- 1:50


for(r in 1:reps){
  slim_code_block <- paste0("slim -d repetition=", r, " model2.1.slim")
  #slim_code_block 
  system(slim_code_block)#call the code block for slim in the terminal via bash, system invokes the OS command specified by command.
}

```



now i want to analyze the data and make one single plot.

```{r}

rm(list=ls(all=TRUE)) #clear R environment

#if the model was run on the cluster - use this command
setwd("/tmp/server_data_model2.1")

slim_out_files <- list.files(pattern = "model2.1_reps_")

length(slim_out_files)
#install.packages("hierfstat")
library("hierfstat")

## make a dataframe to collect results
## each row in the dataframe corresponds to the results from a single SLiM simulation run

row_numbers <- seq(100001, 100139, by = 1)

results_df <- matrix(NA ,nrow = length(slim_out_files), ncol = (length(row_numbers) + 2))
colnames(results_df) <- c(row_numbers,"reps","seed")
length(c(row_numbers,"reps","seed"))
dim(results_df)


for(i in 1:length(slim_out_files)){ 
  
  single_file <- slim_out_files[i] ## this is the name of a single SLiM result file
  get_the_reps <- as.numeric(gsub(".*_reps_([0-9]+)_.*", "\\1", single_file))  ## use gsub() to get the value of 'reps' that we saved in the name of the results file
  get_the_seed <- as.numeric(gsub(".*_seed =|\\.csv","",single_file)) ## and for the seed....
  
  results_df[i,"reps"] <- get_the_reps ## record the value of 'r' for this particular results file
  #.*_reps_ → Matches everything before _reps_
  #([0-9]+) → Captures one or more digits (the reps number)
  #_.* → Matches everything after the number (so it gets removed)
  #"\\1" → Extracts only the number inside ().
  
  results_df[i,"seed"] <- get_the_seed ## record the seed for this particular results file
  
  # Read the CSV file, ensure that it recognizes the header
  df <- read.csv(single_file, header = TRUE) 
  
  # Ensure the column names are correct
  colnames(df) <- c("cycle", "Heterozygosity")  

  # Match cycle numbers to the matrix column names
  matched_indices <- match(as.character(df$cycle), colnames(results_df))

  # Assign heterozygosity values to the correct positions in results_df_nrnr
  results_df[i, matched_indices] <- df$Heterozygosity  

}

# Calculate column means (excluding the last two columns: "reps" and "seed")
column_means <- colMeans(results_df[, 1:(ncol(results_df)-2)], na.rm = TRUE)
# Create a new row with the mean values,  NA for "reps" and "seed" columns
mean_row <- c(column_means, NA, NA)
# Bind the new row to the bottom of results_df_nrnr
results_df <- rbind(results_df, mean_row)
# Assign row names 
rownames(results_df)[nrow(results_df)] <- "Mean"

# Print the final matrix with the mean row
print(results_df)

#write.csv(results_df, "Results_nonrandom_notrelated.csv", row.names = F) #write a file with my results


```




Now do the plot:


```{r}
#install.packages("paletteer")
library(paletteer)


# Extract the heterozygosity values (excluding the last row "Mean" and last two columns "reps" & "seed")
heterozygosity_values <- results_df[1:(nrow(results_df)-1), 1:(ncol(results_df)-2)]
# Extract the mean heterozygosity values from the last row --> Last row, so we can use nrow
mean_values <- results_df[nrow(results_df), 1:(ncol(results_df)-2)]

# Create a vector for the x-values (generations) - this will be used for the x axis of the plot. 
x_values <- 1:139


heterozygosity_only <- heterozygosity_values[, 1:length(x_values)]

#I did this to cut off the last two columns (seed and rep) --> Due to this, i can then apply the filter (next lines) to the matrix. Otherwise it would not match the dimension of cycle_numbers...



#find the highest and lowest heterozygosity values across all three matrices. This will help me with the y axis of the plot i want to generate. 
max_het <- max(heterozygosity_only)
min_het <- min(heterozygosity_only)



#Here i don't have to transpose the data, since each column is a separate data series - this is what matplot expects. 
matplot( x_values, t(heterozygosity_only), # Transpose to align data correctly --> I want to cut the columns not the rows, thats why there is this notation with the 
         type = "p", pch = 16, cex = 0.4,
        col = adjustcolor("#54BFB7FF", alpha.f = 0.02), # Model nrnr: light blue points, rgb(red, green, blue, alpha) enables me to modify the colors a bit better, alpha is the transparacy
        xlab = "Generation", ylab = "Heterozygosity", #name the axis
        ylim = c(0,0.004), main = "Heterozygosity over Generations")  #define the y axis and give the plot a title 
      

#lines( x_values, mean_values, col = "#54BFB7FF", lwd = 2.5)

#Compute median & quartiles (25 and 75 %)
#the 1 means i want to apply the functions over the rows of my table and ignoring NA values (na.rm = TRUE)

med <- apply(t(heterozygosity_only), 1, median, na.rm = TRUE)
p25 <- apply(t(heterozygosity_only), 1, function(x) quantile(x, 0.25, na.rm = TRUE))
p75 <- apply(t(heterozygosity_only), 1, function(x) quantile(x, 0.75, na.rm = TRUE))

lines(x_values, med, col = "tan1", lwd = 2.5, lty = 1)
lines(x_values, p25, col = "#894B33FF", lwd = 2, lty = 3)
lines(x_values, p75, col = "#894B33FF", lwd = 2, lty = 3)

#Add a line to show the bottleneck at ggeneration 35
abline(v = 35,        # x = 36
       lty = 3,       # dashed line 
       col = "red",   #line color
       lwd = 1)       # line width


legend("topright",
       legend = c("all data", "median", "25% & 75% quartile"),
       col    = c("#54BFB7FF",    "tan1",  "#894B33FF"),
       pch    = c(16,            NA,       NA),    # point symbol for first, none for others
       lty    = c(NA,             1,        3),    # no line for first, solid and dashed for rest
       lwd    = c(NA,            2.5,       2),    # no line‐width for first
       pt.cex = c(0.5,           NA,       NA),    # enlarge the point
       cex    = 0.6)


```










