---
title: "R Notebook"
output: html_notebook
---

## Generating Data in SLiM with Parameters from R 

First I have 3 chunks, which will send information to my slim model - With this callback I can make 20 repetitions of my simulation in SLiM and save the output files. I have 3 SLiM models which all show different mateCHoice commands (non-random selecting for non-related mates, non-random selecting for related mates and random mating)

non-random mating selecting for not related mates (nrnr)
```{r eval=FALSE, include=FALSE}

rm(list=ls(all=TRUE)) #clear R environment

options(scipen = 999) #scientific notation penalty, with this R does not use scientific notation for numbers

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory

reps <- 20
repetition <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20) #This tells SliM how many repetitions I want (defined constant in SLiM)


for(r in 1:reps){
  slim_code_block <- paste0("slim -d repetition=", r, " model1_nonrandom_notrelated.slim")
  #slim_code_block - goes through the amount of repetitions I want
  system(slim_code_block)#call the code block for slim in the terminal via bash, system invokes the OS command specified by command.
}

```

non-random mating selecting for related mates (nrr)
```{r eval=FALSE, include=FALSE}

rm(list=ls(all=TRUE)) #clear R environment

options(scipen = 999) #scientific notation penalty, with this R does not use scientific notation for numbers

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory

reps <- 20
repetition <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20) #This tells SliM how many repetitions I want (defined constant in SLiM)


for(r in 1:reps){
  slim_code_block <- paste0("slim -d repetition=", r, " model1_nonrandom_related.slim")
  #slim_code_block - goes through the amount of repetitions I want
  system(slim_code_block)#call the code block for slim in the terminal via bash, system invokes the OS command specified by command.
}

```


random mating (r)
```{r eval=FALSE, include=FALSE}

rm(list=ls(all=TRUE)) #clear R environment

options(scipen = 999) #scientific notation penalty, with this R does not use scientific notation for numbers

setwd("~/Documents/GitHub/2502_emma_bsc") #setting the working directory

reps <- 20
repetition <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20) #This tells SliM how many repetitions I want (defined constant in SLiM)


for(r in 1:reps){
  slim_code_block <- paste0("slim -d repetition=", r, " model1_random.slim")
  #slim_code_block - goes through the amount of repetitions I want
  system(slim_code_block)#call the code block for slim in the terminal via bash, system invokes the OS command specified by command.
}

```



Now i have generated my data and want to analyze them. First I have to order the data to have all the repetitions in one shared matrix. This will make it easier to visualize it in a plot.  

```{r}
#setwd("~/Documents/GitHub/2502_emma_bsc")
setwd("~/Documents/GitHub/2502_emma_bsc/MODEL 1/Model1_100gen_16-1000ind_20reps")
rm(list=ls(all=TRUE)) #clear R environment

#First order the data for the not-random mating, selecting for not-related mates (nrnr)
files_nrnr <- list.files(pattern = "_nonrandom_notrelated_") #list all the files which show this specific pattern in their name. 

length(files_nrnr) #check if it worked, the length should be the same amount as repetitions i defined in the first chunks. 

#install.packages("hierfstat")
library("hierfstat") #This package helps to analyze the data. 

## make a dataframe to collect results
## each row in the dataframe corresponds to the results from a single SLiM simulation run

row_numbers <- seq(100001, 100100, by = 2) #This command will help me name/define the rows. These are the generations from my SLiM output - I had 100 generations and wrote down the heterozygosity for every second generation. 

#row_names <- paste(row_numbers, "reps", "seed", sep = ";")

results_df_nrnr <- matrix(NA ,nrow = (length(row_numbers)), ncol = length(files_nrnr)) #create a matrix with NA, to later store my results. The rows are the different generations and the different columns are the repetitions  I did. 

rownames(results_df_nrnr) <- c(row_numbers) #name the rows of the newly created matrix

dim(results_df_nrnr) #Check if it has the right dimension. 

#for every file in my files_nrnr do...
for(i in 1:length(files_nrnr)){ 
  single_file <- files_nrnr[i] # this is the name of a single SLiM result file
  # Read the CSV file, ensure that it recognizes the header
  df <- read.csv(single_file, header = TRUE) 
   # Ensure the column names are correct
  colnames(df) <- c("cycle", "Heterozygosity")  
  # Match cycle numbers of the single file to the matrix column names
  matched_indices <- match(df$cycle, rownames(results_df_nrnr))

  # Assign heterozygosity values to the correct positions in results_df_nrnr (matrix)
  results_df_nrnr[matched_indices,i] <- df$Heterozygosity
  
}

#calculate the mean over all the rows, so the mean for each generation across the different repetitions. 
mean_nrnr <- rowMeans(results_df_nrnr, na.rm = TRUE)


#now the same for nonrandom mating selecting for relatedness (nrr) 

files_nrr <- list.files(pattern = "_nonrandom_related_")
length(files_nrr)

## make a dataframe to collect results
## each row in the dataframe corresponds to the results from a single SLiM simulation run

row_numbers <- seq(100001, 100100, by = 2)

results_df_nrr <- matrix(NA ,nrow = (length(row_numbers)), ncol = length(files_nrr))
rownames(results_df_nrr) <- c(row_numbers)

dim(results_df_nrr)

for(i in 1:length(files_nrr)){ 
  single_file <- files_nrr[i] # this is the name of a single SLiM result file
  # Read the CSV file, ensure that it recognizes the header
  df <- read.csv(single_file, header = TRUE) 
   # Ensure the column names are correct
  colnames(df) <- c("cycle", "Heterozygosity")  
  # Match cycle numbers to the matrix column names
  matched_indices <- match(df$cycle, rownames(results_df_nrr))

  # Assign heterozygosity values to the correct positions in results_df_nrr
  results_df_nrr[matched_indices,i] <- df$Heterozygosity
  
}

mean_nrr <- rowMeans(results_df_nrr, na.rm = TRUE)


#and for random

files_r <- list.files(pattern = "_random_")
length(files_r)

## make a dataframe to collect results
## each row in the dataframe corresponds to the results from a single SLiM simulation run

row_numbers <- seq(100001, 100100, by = 2)
#row_names <- paste(row_numbers, "reps", "seed", sep = ";")

results_df_r <- matrix(NA ,nrow = (length(row_numbers)), ncol = length(files_r))
rownames(results_df_r) <- c(row_numbers)

dim(results_df_r)

for(i in 1:length(files_r)){ 
  single_file <- files_r[i] ## this is the name of a single SLiM result file
  # Read the CSV file, ensure that it recognizes the header
  df <- read.csv(single_file, header = TRUE) 
   # Ensure the column names are correct
  colnames(df) <- c("cycle", "Heterozygosity")  
  # Match cycle numbers to the matrix column names
  matched_indices <- match(df$cycle, rownames(results_df_r))

  # Assign heterozygosity values to the correct positions in results_df_r
  results_df_r[matched_indices,i] <- df$Heterozygosity
  
}

mean_r <- rowMeans(results_df_r, na.rm = TRUE)


```

```{r}
setwd("~/Documents/GitHub/2502_emma_bsc/MODEL 1/Model1_100gen_16-1000ind_20reps") #setting the working directory


#install.packages("paletteer")
library(paletteer)


#find the highest and lowest heterozygosity values across all three matrices. This will help me with the y axis of the plot i want to generate. 
max_het <- max(results_df_nrnr, results_df_nrr, results_df_r)
min_het <- min(results_df_nrnr, results_df_nrr, results_df_r)

# Create a vector for the x-values (generations) - this will be used for the x axis of the plot. 
x_values <- seq(2, 100, by = 2)

#Here i don't have to transpose the data, since each column is a separate data series - this is what matplot expects. 
matplot( x_values, results_df_nrnr, type = "p", pch = 16, 
        col = adjustcolor("#54BFB7FF", alpha.f = 0.1), # Model nrnr: light blue points, rgb(red, green, blue, alpha) enables me to modify the colors a bit better, alpha is the transparacy
        xlab = "Generation", ylab = "Heterozygosity", #name the axis
        ylim = c(min_het, max_het), main = "Heterozygosity over Generations")  #define the y axis and give the plot a title 

#Now I generated the plot and can just add the data of the other models (nrr and r)
# Model nrr: light green points
matpoints(x_values, results_df_nrr, pch = 16, col = adjustcolor("#894B33FF", alpha.f = 0.1)) #adding points to the existing plot from above

# Model r: light purple points
matpoints( x_values, results_df_r, pch = 16, col = adjustcolor("tan1", alpha.f = 0.1)) #again

# adding Mean relatedness lines to the plot
lines( x_values, mean_nrnr, col = "#54BFB7FF", lwd = 2)
lines( x_values, mean_nrr, col = "#894B33FF", lwd = 2)
lines( x_values, mean_r, col = "tan1", lwd = 2)

#create a legend for my plot
legend("bottomleft", legend=c("non-random non-related", "non-rendom related", "random"), col=c("#54BFB7FF", "#894B33FF", "tan1"), lty=1, cex = 0.7)
#cex = 0.7 makes the legend smaller, lty is for the lines


```
